Body size patterns

### 2019-06-28 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
new models

system.time(
mod0_YDCcr_C_CSr_Tr_Fr_TFr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_CR_Cr_CS_Tr_Fr_TFr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_CR_Cr_CS_Tr_Fr_TFr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_CR_Cr_CSr_T_F_TF <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr'), 
               data = dataIn, 
               method = "REML")
)

@@

system.time(
mod0_CR_Cr_CSr <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered),  # this is conspecific cS only
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_CR_Cr_Tr_Fr_TFr <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_CR_CSr_Tr_Fr_TFr <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_CSr_Tr_Fr_TFr <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered), 
               data = dataIn, 
               method = "REML")
)
########

system.time(
mod0_CR_Cr_CS_T_F_TF <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr'), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_CR_Cr_CS_Tr_F_TF <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_C_Cr_CS_Tr_F_TF <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverOrdered) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn), 
               data = dataIn, 
               method = "REML")
)
####
system.time(
mod0_C_Cr_CS_T_F_TF <- bam(observedLength ~ cohort * factor(river) * factor(salmon01) + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
              # s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr'), 
               data = dataIn, 
               method = "REML")
)

tmp <- 
bam(observedLength ~ factor(river) * factor(season) + factor(salmon01) + 
      #factor(river) *  + meanTemperatureScaledBySeason * meanFlowScaledBySeason * factor(season) + cohortStrengthScaled +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
              # s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
                s(cohort, bs = 'cr', k = kIn, by = riverOrdered) + 
                s(cohortStrengthScaled, bs = 'cr', k = kIn) +  # this is conspecific cS only
                s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = season) +
                s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = season) +
                s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = season), 
               data = dataIn, 
               method = "REML")
summary(tmp)
plot(tmp, scale = 0, shade = TRUE)

AIC(mod0, mod0_CR, mod0_CR_Cr_CSr_Tr_Fr_TFr,mod0_CR_C_CSr_Tr_Fr_TFr,mod0_CR_Cr_CS_Tr_Fr_TFr,mod0_CR_Cr_CSr_T_F_TF,
    mod0_CR_Cr_CSr,mod0_CR_Cr_Tr_Fr_TFr,mod0_CR_CSr_Tr_Fr_TFr,mod0_CSr_Tr_Fr_TFr,mod0_CR_Cr_CS_T_F_TF,mod0_CR_Cr_CS_Tr_F_TF,mod0_C_Cr_CS_Tr_F_TF,mod0_C_Cr_CS_T_F_TF) %>% rownames_to_column() %>% arrange(AIC)

summary(mod0_CR_Cr_CS_T_F_TF)
gam.check(mod0_CR_Cr_CS_T_F_TF)
plot(mod0_CR_Cr_CS_T_F_TF, scale = 0, pages=8)

@@

```{r}
predGrid <- expand.grid(cohort = unique(mod$model$cohort),
                        river = unique(mod$model$river),
                        ydayCumul = seq(250,1000, by = 30))

pred <- predict(mod, newdata = predGrid)#, type = 'lpmatrix')

ggplot(predGrid2, aes(ydayCumul, pred)) +
  geom_point() + 
  geom_line() +
  facet_grid(river ~ cohort) +
  ggtitle('mod')

##########################################################
# by cohortSpeciesRiver

system.time(
modCR <- bam(observedLength ~ cohort * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered), 
             data = dataIn, method = "REML")
)
summary(modCR)
#plot(modCR, pages = 8, scale = 0)

predGridCR <- expand.grid(cohortRiverOrdered = unique(modCR$model$cohortRiverOrdered),
                           ydayCumul = seq(250,1000, by = 30)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
                           riverOrdered = substring(cohortRiverOrdered,6),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           river = riverOrdered)

predCR <- predict(modCR, newdata = predGridCR)#, type = 'lpmatrix')

predGrid2CR <- bind_cols(predGridCR, as.data.frame(predCR))

ggplot(predGrid2CR, aes(ydayCumul, predCR)) +
  geom_point() + 
  geom_line() +
  ylim(50,210) +
  facet_grid(river ~ cohort)+
  ggtitle('modCR')

##########################################################
# by cohortSpeciesRiver with s(cohort)

system.time(
modCR_c <- bam(observedLength ~ cohort * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn), 
             data = dataIn, method = "REML")
)
summary(modCR_c)
#plot(modCR_c, pages = 8, scale = 0)

predGridCR_c <- expand.grid(cohortRiverOrdered = unique(modCR$model$cohortRiverOrdered),
                           ydayCumul = seq(250,1000, by = 30)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
                           riverOrdered = substring(cohortRiverOrdered,6),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           river = riverOrdered)

predCR_c <- predict(modCR_c, newdata = predGridCR_c)#, type = 'lpmatrix')

predGrid2CR_c <- bind_cols(predGridCR_c, as.data.frame(predCR_c))

ggplot(predGrid2CR_c, aes(ydayCumul, predCR_c, group = cohort, color = cohort)) +
  geom_point() + 
  geom_line() +
  #ylim(50,210) +
  facet_wrap(~river)+
  ggtitle('modCR_c')

##########################################################
# by cohortRiver - w ind RE

# need more memory
# system.time(
# modCR_Ind <- bam(observedLength ~ cohort * river +
#                s(ydayCumul, bs = 'cr') +
#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
#                s(ydayCumul, tagIndexFactor, bs="fs", m=1, k = kIn),
#              data = dataIn, method = "REML")
# )
# summary(modCR_Ind)
# plot(modCR_Ind, pages = 8, scale = 0)

##########################################################
# by cohortSpeciesRiver, additive in s()

system.time(
modC_R <- bam(observedLength ~ cohort * river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered), 
             data = dataIn, method = "REML")
)
summary(modC_R)
#plot(modC_R, pages = 4, scale = 0)

predGridC_R <- expand.grid(cohortRiverOrdered = unique(modCR$model$cohortRiverOrdered),
                           ydayCumul = seq(250,1000, by = 30)) %>%
                    mutate(cohortOrdered = as.numeric(substr(cohortRiverOrdered,1,4)),
                           riverOrdered = substring(cohortRiverOrdered,6),
                           cohortRiverOrdered = paste(cohortOrdered, riverOrdered, sep = '_'),
                           cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortOrdered,
                           river = riverOrdered)

predC_R <- predict(modC_R, newdata = predGridC_R)#, type = 'lpmatrix')

predGrid2C_R <- bind_cols(predGridC_R,as.data.frame(predC_R))

ggplot(predGrid2C_R, aes(ydayCumul,predC_R)) +
  geom_point() + 
  geom_line() +
  ylim(50,210) +
  facet_grid(river ~ cohort)+
  ggtitle('modC_R')

##########################

modC <- bam(observedLength ~ cohort + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered), 
             data = vB %>% filter(cohort %in% 2000:2013,species %in% c('bkt')), method = "REML")

modR <- bam(observedLength ~ river + 
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered), 
             data = dataIn, method = "REML")

##########################################
# add in salmon01 and cohortStrengthScaled

system.time(
modCR_S_c <- bam(observedLength ~ cohort * river + factor(salmon01) +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn), 
             data = dataIn, method = "REML")
)
system.time(
modCR_c_cS <- bam(observedLength ~ cohort * river +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn),  # this is conspecific cS only
             data = dataIn, method = "REML")
)

system.time(
modCR_S_c_cS <- bam(observedLength ~ cohort * river + factor(salmon01) +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn),  # this is conspecific cS only
             data = dataIn, method = "REML")
)
summary(modCR_S_c_cS)

system.time(
modCR_SR_c_cS <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn),  # this is conspecific cS only
             data = dataIn, method = "REML")
)
summary(modCR_SR_c_cS)

system.time(
modCR_SR_c_cSR <- bam(observedLength ~ cohort * river + factor(salmon01) * river +
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn,  by = riverOrdered) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverOrdered) +
               s(cohort, bs = 'cr', k = kIn) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverOrdered),  # this is conspecific cS only
             data = dataIn, method = "REML")
)
summary(modCR_SR_c_cSR)
#plot(modCR_S_c_cS, scale = 0, pages=8)
#plotGamVarBySpecies(var = 's(cohortStrengthScaled)')

>>>>

@@

system.time(
mod0_YDCcr_Cr_CSr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_YDCcr_Cr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               #s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_YDCcr_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
              # s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_YDCcr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
             #  s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
             #  s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
               data = dataIn, 
               method = "REML")
)
########

AIC(mod0_YDCcr, mod0_YDCc_r, mod0_YDCcr_Cr_CSr_Tr_Fr_TFr, mod0_YDCcr_Cr_CSr_Ts_Fs_TFs,mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_CSr,mod0_YDCcr_C_CS_Tsr_Fsr_TFsr,mod0_YDCcr_Cr_Tsr_Fsr_TFsr,mod0_YDCcr_CSr_Tsr_Fsr_TFsr, mod0_YDCcr_Tsr_Fsr_TFsr) %>% rownames_to_column() %>% arrange(AIC)

#                           rowname        df      AIC
# 1  mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr 277.67003 401304.4
# 2   mod0_YDCcr_C_CSr_Tsr_Fsr_TFsr 281.22421 401407.3
# 3   mod0_YDCcr_Cr_CS_Tsr_Fsr_TFsr 277.09498 401772.0
# 4     mod0_YDCcr_CSr_Tsr_Fsr_TFsr 279.84632 402037.9
# 5    mod0_YDCcr_C_CS_Tsr_Fsr_TFsr 269.69828 402073.1
# 6     mod0_YDCcr_Cr_CSr_Tr_Fr_TFr 207.97536 402187.1
# 7     mod0_YDCcr_Cr_CSr_Ts_Fs_TFs 190.64645 402464.7
# 8      mod0_YDCcr_Cr_Tsr_Fsr_TFsr 289.33427 403529.7
# 9         mod0_YDCcr_Tsr_Fsr_TFsr 283.93242 404139.1
# 10              mod0_YDCcr_Cr_CSr 184.24038 413605.9
# 11                     mod0_YDCcr 168.77693 420503.0
# 12                    mod0_YDCc_r  71.99108 422677.4

summary(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr) 
plot(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr, pages = 8, shade = TRUE, scale = 0)

@@

```{r predictions}
# BKT
# system.time(
# mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
#                s(ydayCumul, bs = 'cr') +
#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
#                s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
#                s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
#                s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
#                s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
#                s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor) +
#                s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = seasonRiverFactor), 
#                data = dataIn, 
#                method = "REML")
# )

predGridBKT <- expand.grid(cohortRiverFactor = unique(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr$model$cohortRiverFactor),
                           seasonRiverFactor = unique(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr$model$seasonRiverFactor),
                           ydayCumul = seq(250,1000, by = 30),
                           salmon01 = c(0,1),
                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
                           meanFlowScaledBySeason = 0) %>% #seq(-1.5,1.5, by = 0.5)) %>%
                          # meanTemperatureFlowScaledBySeason = seq(-1.5,1.5, by = 0.5)) %>%
                    mutate(cohortFactor = as.numeric(substr(cohortRiverFactor,1,4)),
                           riverFactor = substring(cohortRiverFactor,6),
                           season = substring(seasonRiverFactor,1),
                           #cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
                           #seasonRiverFactor = paste(season, riverFactor, sep = '_'),
                           #cohortCentered = cohortOrdered - cohortOffset,
                           cohort = cohortFactor,
                           #river = riverFactor,
                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)

ydayCumulSeason <- distinct(dataIn, ydayCumul,season,age) %>% arrange(ydayCumul) %>% mutate(age_season = paste(age,season,sep = '_')) %>%
  group_by(age_season) %>% summarise(age=min(age),season=min(season),minY=min(ydayCumul),maxY=max(ydayCumul))
#38 obear fish were captured in january, but get season 4 and are age 1, so have very small ydayCumul -need to manually reset 
#dataIn %>% filter(age==1,season==4,ydayCumul==367) %>% select(tag,cohort,detectionDate,year,ydayCumul,yday,river,age,season)
ydayCumulSeason$minY <- ifelse(ydayCumulSeason$age_season == '1_4',670,ydayCumulSeason$minY)
ydayCumulSeason$minY <- ifelse(ydayCumulSeason$age_season == '2_4',1035,ydayCumulSeason$minY)
ydayCumulSeason$minY <- ifelse(ydayCumulSeason$age_season == '0_3',0,ydayCumulSeason$minY)
ydayCumulSeason %>% mutate(minY2 = lead(minY))

yDayAll <- data.frame(y = 250:1000)
# pick up here
#### need to do something like this  left_join(yDayAll, ydayCumulSeason, by = c('yDay' = 'minY'))

predGridBKT <- expand.grid(riverFactor = unique(dataIn$riverFactor), 
                           cohortFactor = unique(dataIn$cohort),
                           ydayCumul = seq(250,1000, by = 30),
                           salmon01 = c(0,1),
                           cohortStrengthScaled = 0, #seq(-1.5,1.5, by = 0.5),
                           meanTemperatureScaledBySeason = 0, #seq(-1.5,1.5, by = 0.5),
                           meanFlowScaledBySeason = 0) %>%
                    left_join(., ydayCumulSeason) %>%
                    mutate(cohortRiverFactor = paste(cohortFactor, riverFactor, sep = '_'),
                           seasonRiverFactor = paste(season, riverFactor, sep = '_'),
                           cohort = cohortFactor,
                           #river = riverFactor,
                           meanTemperatureFlowScaledBySeason = meanTemperatureScaledBySeason * meanFlowScaledBySeason)

predBKT <- predict(mod0_YDCcr_Cr_CSr_Tsr_Fsr_TFsr, newdata = predGridBKT)#, type = 'lpmatrix')

ggplot(predGrid2BKT %>% filter(salmon01 == 0),

       aes(ydayCumul, pred, group = cohort, color = cohort)) +
  #geom_point() + 
  geom_vline(xintercept = c(365,365*2), color = 'grey60', linetype = 2) +
  geom_line(size = 1.1) +
  labs(y = 'Predicted body size (mm)',
       x = 'Days since age-0 Jan 1',
       color = 'Cohort') +
  scale_color_viridis_c(breaks = c(2000,2004,2008,2012), labels = c(2000,2004,2008,2012)) +
  theme_publication() +
  theme(legend.position = "right", 
        legend.direction = "vertical",
        legend.title = element_text(face = "plain")) +
  guides(color = guide_colorbar(barheight = 10,
                                barwidth = 1.5,
                                ticks = FALSE,
                                nbin = 100)) +
#  guide_legend(title = 'Cohort') +
  facet_grid(~riverFactor, scales = 'free_y')

```{r}
#earlier runs before setting fixed effect first. kept forf reference
# BKT 
#             rowname        df      AIC
# 1     modCR_SR_c_cS 162.34756 414267.2
# 2      modCR_S_c_cS 163.43154 414602.7
# 3        modCR_c_cS 161.63691 414750.3
# 4  modCR_SR_c_cSR3a  88.21646 415388.9
# 5  modCR_SR_c_cSR3b 167.32968 416062.3
# 6    modCR_SR_c_cSR 163.93920 416587.5
# 7   modCR_SR_c_cSR3 173.36884 416611.4
# 8         modCR_S_c 157.23995 418645.9
# 9           modCR_c 158.50186 418665.1
# 10            modCR 154.36186 419135.1
# 11           modC_R  63.10879 421258.1
# 12              mod  17.99204 427309.6
# 13             modC  48.73277 431440.2
# 14             modR  22.80483 432214.7
